# Preprocessing
## library A
### Qualty Control, hashing demultiplexing

```{r message=FALSE, warning=FALSE, eval=T, fig.height=6, fig.width=16}

Lib <- "A"
patient <-  c("C686518","C686518","H230","H230")
tags <-  c("C686518","C686518.BMP9","H230","H230.BMP9")

data <- Read10X(paste0(workdir,"/data/Lib_",Lib))
hto <- Read10X(paste0(workdir, "/data/hto/Lib_",Lib,"_umi_count"), gene.column=1)
hto = hto[1:4,]
hto[1:4,1:5]
Matrix::rowSums(hto)

test="-1"
colnames(hto) = paste(colnames(hto),test,sep = "")
common <- intersect(colnames(data), colnames(hto))
data <- data[,common]
hto <- hto[,common]
rownames(hto) <- tags

hashing <- CreateSeuratObject(data)
hashing[["HTO"]] <- CreateAssayObject(hto)
hashing <- NormalizeData(object = hashing, assay = "HTO", normalization.method = "CLR")
hashing <- HTODemux(hashing)

mito.genes <- grep(pattern = "^MT-", x = rownames(hashing@assays$RNA), value = TRUE)
ribo.genes <- grep(pattern = "^RP[SL]", x = rownames(hashing@assays$RNA), value = TRUE)
dropouts <- Matrix::colSums(hashing@assays$RNA@data == 0)/nrow(hashing@assays$RNA)
percent.mito <- Matrix::colSums(hashing@assays$RNA[mito.genes, ])/Matrix::colSums(hashing@assays$RNA)
percent.ribo <- Matrix::colSums(hashing@assays$RNA[ribo.genes, ])/Matrix::colSums(hashing@assays$RNA)
hashing[['percent.mito']] <- percent.mito
hashing[['percent.ribo']] <- percent.ribo
hashing[['dropouts']] <- dropouts
VlnPlot(hashing, features = c("nFeature_RNA", "nCount_RNA","dropouts","percent.ribo","percent.mito"), ncol=5)
dim(hashing@assays$RNA)

hashing <- subset(hashing, subset = percent.mito < 0.15)
dim(hashing@assays$RNA)
VlnPlot(hashing, features = c("nFeature_RNA", "nCount_RNA","dropouts","percent.ribo","percent.mito"), ncol=5)

print(table(hashing@meta.data$HTO_classification.global))
hashing <- SetIdent(hashing, value = "HTO_maxID")

HTOHeatmap(hashing)

```

#### Singlet only

```{r message=FALSE, warning=FALSE, eval=T, fig.height=6, fig.width=12 }

hashing <- SetIdent(hashing, value = "HTO_classification.global")
singlet <- subset(hashing, idents = "Singlet")
table(singlet@meta.data$HTO_maxID)

```


#### Tag analysis

```{r message=FALSE, warning=FALSE, eval=T, fig.height=6, fig.width=8 }

### Tag1 analysis

tag1 <- subset(singlet, subset = HTO_maxID == tags[1])
tag1@meta.data$library <- Lib
tag1@meta.data$tag <- paste0(Lib,"1")
tag1@meta.data$BMP9 <- "0 ng/ml"
tag1@meta.data$phenotype <- "CTRL"
tag1@meta.data$patient <- patient[1]
tag1@meta.data$sample <- tags[1]

tag1 <- SCTransform(object = tag1, verbose = FALSE)
tag1 <- RunPCA(tag1, verbose = FALSE)
tag1 <- RunUMAP(object = tag1, dims = 1:30)
tag1 <- FindNeighbors(object = tag1, do.plot=TRUE, dims = 1:30)
tag1 <- FindClusters(object = tag1, resolution=0.4)
DimPlot(object = tag1, label=TRUE)

tag1 <- RenameCells(tag1, add.cell.id = paste0(Lib,"1"))
saveRDS(tag1, paste0(outdir,"/samples/",Lib,"1",".rds"))

### Tag2 analysis

tag2 <- subset(singlet, subset = HTO_maxID == tags[2])
tag2@meta.data$library <- Lib
tag2@meta.data$tag <- paste0(Lib,"2")
tag2@meta.data$BMP9 <- "10 ng/ml"
tag2@meta.data$phenotype <- "CTRL.BMP9"
tag2@meta.data$patient <- patient[2]
tag2@meta.data$sample <- tags[2]

tag2 <- SCTransform(object = tag2, verbose = FALSE)
tag2 <- RunPCA(tag2, verbose = FALSE)
tag2 <- RunUMAP(object = tag2, dims = 1:30)
tag2 <- FindNeighbors(object = tag2, do.plot=TRUE, dims = 1:30)
tag2 <- FindClusters(object = tag2, resolution=0.4)
DimPlot(object = tag2, label=TRUE)

tag2 <- RenameCells(tag2, add.cell.id = paste0(Lib,"2"))
saveRDS(tag2, paste0(outdir,"/samples/",Lib,"2",".rds"))

### Tag3 analysis

tag3 <- subset(singlet, subset = HTO_maxID == tags[3])
tag3@meta.data$library <- Lib
tag3@meta.data$tag <- paste0(Lib,"3")
tag3@meta.data$BMP9 <- "0 ng/ml"
tag3@meta.data$phenotype <- "HTAP"
tag3@meta.data$patient <- patient[3]
tag3@meta.data$sample <- tags[3]

tag3 <- SCTransform(object = tag3, verbose = FALSE)
tag3 <- RunPCA(tag3, verbose = FALSE)
tag3 <- RunUMAP(object = tag3, dims = 1:30)
tag3 <- FindNeighbors(object = tag3, do.plot=TRUE, dims = 1:30)
tag3 <- FindClusters(object = tag3, resolution=0.4)
invisible(gc())

DimPlot(object = tag3, label=TRUE)

tag3 <- RenameCells(tag3, add.cell.id = paste0(Lib,"3"))
saveRDS(tag3, paste0(outdir,"/samples/",Lib,"3",".rds"))

### Tag4 analysis

tag4 <- subset(singlet, subset = HTO_maxID == tags[4])
tag4@meta.data$library <- Lib
tag4@meta.data$tag <- paste0(Lib,"4")
tag4@meta.data$BMP9 <- "10 ng/ml"
tag4@meta.data$phenotype <- "HTAP.BMP9"
tag4@meta.data$patient <- patient[4]
tag4@meta.data$sample <- tags[4]

tag4 <- SCTransform(object = tag4, verbose = FALSE)
tag4 <- RunPCA(tag4, verbose = FALSE)
tag4 <- RunUMAP(object = tag4, dims = 1:30)
tag4 <- FindNeighbors(object = tag4, do.plot=TRUE, dims = 1:30)
tag4 <- FindClusters(object = tag4, resolution=0.4)
invisible(gc())

DimPlot(object = tag4, label=TRUE)

tag4 <- RenameCells(tag4, add.cell.id = paste0(Lib,"4"))
saveRDS(tag4, paste0(outdir,"/samples/",Lib,"4",".rds"))

```


## library B
### Qualty Control, hashing demultiplexing

```{r message=FALSE, warning=FALSE, eval=T, fig.height=6, fig.width=16}

Lib <- "B"
patient <-  c("C681725","C681725","H231","H231")
tags <-  c("C681725","C681725.BMP9","H231","H231.BMP9")

data <- Read10X(paste0(workdir,"/data/Lib_",Lib))
hto <- Read10X(paste0(workdir, "/data/hto/Lib_",Lib,"_umi_count"), gene.column=1)
hto = hto[1:4,]
hto[1:4,1:5]
Matrix::rowSums(hto)

test="-1"
colnames(hto) = paste(colnames(hto),test,sep = "")
common <- intersect(colnames(data), colnames(hto))
data <- data[,common]
hto <- hto[,common]
rownames(hto) <- tags

hashing <- CreateSeuratObject(data)
hashing[["HTO"]] <- CreateAssayObject(hto)
hashing <- NormalizeData(object = hashing, assay = "HTO", normalization.method = "CLR")
hashing <- HTODemux(hashing)

mito.genes <- grep(pattern = "^MT-", x = rownames(hashing@assays$RNA), value = TRUE)
ribo.genes <- grep(pattern = "^RP[SL]", x = rownames(hashing@assays$RNA), value = TRUE)
dropouts <- Matrix::colSums(hashing@assays$RNA@data == 0)/nrow(hashing@assays$RNA)
percent.mito <- Matrix::colSums(hashing@assays$RNA[mito.genes, ])/Matrix::colSums(hashing@assays$RNA)
percent.ribo <- Matrix::colSums(hashing@assays$RNA[ribo.genes, ])/Matrix::colSums(hashing@assays$RNA)
hashing[['percent.mito']] <- percent.mito
hashing[['percent.ribo']] <- percent.ribo
hashing[['dropouts']] <- dropouts
VlnPlot(hashing, features = c("nFeature_RNA", "nCount_RNA","dropouts","percent.ribo","percent.mito"), ncol=5)
dim(hashing@assays$RNA)

hashing <- subset(hashing, subset = percent.mito < 0.15)
dim(hashing@assays$RNA)
VlnPlot(hashing, features = c("nFeature_RNA", "nCount_RNA","dropouts","percent.ribo","percent.mito"), ncol=5)

print(table(hashing@meta.data$HTO_classification.global))
hashing <- SetIdent(hashing, value = "HTO_maxID")

HTOHeatmap(hashing)

```

#### Singlet only

```{r message=FALSE, warning=FALSE, eval=T, fig.height=6, fig.width=12 }

hashing <- SetIdent(hashing, value = "HTO_classification.global")
singlet <- subset(hashing, idents = "Singlet")
table(singlet@meta.data$HTO_maxID)

```

#### Tag analysis

```{r message=FALSE, warning=FALSE, eval=T, fig.height=6, fig.width=8 }

### Tag1 analysis

tag1 <- subset(singlet, subset = HTO_maxID == tags[1])
tag1@meta.data$library <- Lib
tag1@meta.data$tag <- paste0(Lib,"1")
tag1@meta.data$BMP9 <- "0 ng/ml"
tag1@meta.data$phenotype <- "CTRL"
tag1@meta.data$patient <- patient[1]
tag1@meta.data$sample <- tags[1]

tag1 <- SCTransform(object = tag1, verbose = FALSE)
tag1 <- RunPCA(tag1, verbose = FALSE)
tag1 <- RunUMAP(object = tag1, dims = 1:30)
tag1 <- FindNeighbors(object = tag1, do.plot=TRUE, dims = 1:30)
tag1 <- FindClusters(object = tag1, resolution=0.4)
invisible(gc())

DimPlot(object = tag1, label=TRUE)

tag1 <- RenameCells(tag1, add.cell.id = paste0(Lib,"1"))
saveRDS(tag1, paste0(outdir,"/samples/",Lib,"1",".rds"))

### Tag2 analysis

tag2 <- subset(singlet, subset = HTO_maxID == tags[2])
tag2@meta.data$library <- Lib
tag2@meta.data$tag <- paste0(Lib,"2")
tag2@meta.data$BMP9 <- "10 ng/ml"
tag2@meta.data$phenotype <- "CTRL.BMP9"
tag2@meta.data$patient <- patient[2]
tag2@meta.data$sample <- tags[2]

tag2 <- SCTransform(object = tag2, verbose = FALSE)
tag2 <- RunPCA(tag2, verbose = FALSE)
tag2 <- RunUMAP(object = tag2, dims = 1:30)
tag2 <- FindNeighbors(object = tag2, do.plot=TRUE, dims = 1:30)
tag2 <- FindClusters(object = tag2, resolution=0.4)
invisible(gc())
DimPlot(object = tag2, label=TRUE)

tag2 <- RenameCells(tag2, add.cell.id = paste0(Lib,"2"))
saveRDS(tag2, paste0(outdir,"/samples/",Lib,"2",".rds"))

### Tag3 analysis

tag3 <- subset(singlet, subset = HTO_maxID == tags[3])
tag3@meta.data$library <- Lib
tag3@meta.data$tag <- paste0(Lib,"3")
tag3@meta.data$BMP9 <- "0 ng/ml"
tag3@meta.data$phenotype <- "HTAP"
tag3@meta.data$patient <- patient[3]
tag3@meta.data$sample <- tags[3]

tag3 <- SCTransform(object = tag3, verbose = FALSE)
tag3 <- RunPCA(tag3, verbose = FALSE)
tag3 <- RunUMAP(object = tag3, dims = 1:30)
tag3 <- FindNeighbors(object = tag3, do.plot=TRUE, dims = 1:30)
tag3 <- FindClusters(object = tag3, resolution=0.4)
invisible(gc())

DimPlot(object = tag3, label=TRUE)

tag3 <- RenameCells(tag3, add.cell.id = paste0(Lib,"3"))
saveRDS(tag3, paste0(outdir,"/samples/",Lib,"3",".rds"))

### Tag4 analysis

tag4 <- subset(singlet, subset = HTO_maxID == tags[4])
tag4@meta.data$library <- Lib
tag4@meta.data$tag <- paste0(Lib,"4")
tag4@meta.data$BMP9 <- "10 ng/ml"
tag4@meta.data$phenotype <- "HTAP.BMP9"
tag4@meta.data$patient <- patient[4]
tag4@meta.data$sample <- tags[4]

tag4 <- SCTransform(object = tag4, verbose = FALSE)
tag4 <- RunPCA(tag4, verbose = FALSE)
tag4 <- RunUMAP(object = tag4, dims = 1:30)
tag4 <- FindNeighbors(object = tag4, do.plot=TRUE, dims = 1:30)
tag4 <- FindClusters(object = tag4, resolution=0.4)
invisible(gc())

DimPlot(object = tag4, label=TRUE)

tag4 <- RenameCells(tag4, add.cell.id = paste0(Lib,"4"))
saveRDS(tag4, paste0(outdir,"/samples/",Lib,"4",".rds"))

```


## library C
### Qualty Control, hashing demultiplexing

```{r message=FALSE, warning=FALSE, eval=T, fig.height=6, fig.width=16}

Lib <- "C"
patient <-  c("C681086","C681086","H247","H247")
tags <-  c("C681086","C681086.BMP9","H247","H247.BMP9")

data <- Read10X(paste0(workdir,"/data/Lib_",Lib))
hto <- Read10X(paste0(workdir, "/data/hto/Lib_",Lib,"_umi_count"), gene.column=1)
hto = hto[1:4,]
hto[1:4,1:5]
Matrix::rowSums(hto)

test="-1"
colnames(hto) = paste(colnames(hto),test,sep = "")
common <- intersect(colnames(data), colnames(hto))
data <- data[,common]
hto <- hto[,common]
rownames(hto) <- tags

hashing <- CreateSeuratObject(data)
hashing[["HTO"]] <- CreateAssayObject(hto)
hashing <- NormalizeData(object = hashing, assay = "HTO", normalization.method = "CLR")
hashing <- HTODemux(hashing)

mito.genes <- grep(pattern = "^MT-", x = rownames(hashing@assays$RNA), value = TRUE)
ribo.genes <- grep(pattern = "^RP[SL]", x = rownames(hashing@assays$RNA), value = TRUE)
dropouts <- Matrix::colSums(hashing@assays$RNA@data == 0)/nrow(hashing@assays$RNA)
percent.mito <- Matrix::colSums(hashing@assays$RNA[mito.genes, ])/Matrix::colSums(hashing@assays$RNA)
percent.ribo <- Matrix::colSums(hashing@assays$RNA[ribo.genes, ])/Matrix::colSums(hashing@assays$RNA)
hashing[['percent.mito']] <- percent.mito
hashing[['percent.ribo']] <- percent.ribo
hashing[['dropouts']] <- dropouts
VlnPlot(hashing, features = c("nFeature_RNA", "nCount_RNA","dropouts","percent.ribo","percent.mito"), ncol=5)
dim(hashing@assays$RNA)

hashing <- subset(hashing, subset = percent.mito < 0.15)
dim(hashing@assays$RNA)
VlnPlot(hashing, features = c("nFeature_RNA", "nCount_RNA","dropouts","percent.ribo","percent.mito"), ncol=5)

print(table(hashing@meta.data$HTO_classification.global))
hashing <- SetIdent(hashing, value = "HTO_maxID")

HTOHeatmap(hashing)

```

#### Singlet only

```{r message=FALSE, warning=FALSE, eval=T, fig.height=6, fig.width=12 }

hashing <- SetIdent(hashing, value = "HTO_classification.global")
singlet <- subset(hashing, idents = "Singlet")
table(singlet@meta.data$HTO_maxID)

```


#### Tag analysis

```{r message=FALSE, warning=FALSE, eval=T, fig.height=6, fig.width=8 }

### Tag1 analysis

tag1 <- subset(singlet, subset = HTO_maxID == tags[1])
tag1@meta.data$library <- Lib
tag1@meta.data$tag <- paste0(Lib,"1")
tag1@meta.data$BMP9 <- "0 ng/ml"
tag1@meta.data$phenotype <- "CTRL"
tag1@meta.data$patient <- patient[1]
tag1@meta.data$sample <- tags[1]

tag1 <- SCTransform(object = tag1, verbose = FALSE)
tag1 <- RunPCA(tag1, verbose = FALSE)
tag1 <- RunUMAP(object = tag1, dims = 1:30)
tag1 <- FindNeighbors(object = tag1, do.plot=TRUE, dims = 1:30)
tag1 <- FindClusters(object = tag1, resolution=0.4)
invisible(gc())

DimPlot(object = tag1, label=TRUE)

tag1 <- RenameCells(tag1, add.cell.id = paste0(Lib,"1"))
saveRDS(tag1, paste0(outdir,"/samples/",Lib,"1",".rds"))

### Tag2 analysis

tag2 <- subset(singlet, subset = HTO_maxID == tags[2])
tag2@meta.data$library <- Lib
tag2@meta.data$tag <- paste0(Lib,"2")
tag2@meta.data$BMP9 <- "10 ng/ml"
tag2@meta.data$phenotype <- "CTRL.BMP9"
tag2@meta.data$patient <- patient[2]
tag2@meta.data$sample <- tags[2]

tag2 <- SCTransform(object = tag2, verbose = FALSE)
tag2 <- RunPCA(tag2, verbose = FALSE)
tag2 <- RunUMAP(object = tag2, dims = 1:30)
tag2 <- FindNeighbors(object = tag2, do.plot=TRUE, dims = 1:30)
tag2 <- FindClusters(object = tag2, resolution=0.4)
invisible(gc())

DimPlot(object = tag2, label=TRUE)

tag2 <- RenameCells(tag2, add.cell.id = paste0(Lib,"2"))
saveRDS(tag2, paste0(outdir,"/samples/",Lib,"2",".rds"))

### Tag3 analysis

tag3 <- subset(singlet, subset = HTO_maxID == tags[3])
tag3@meta.data$library <- Lib
tag3@meta.data$tag <- paste0(Lib,"3")
tag3@meta.data$BMP9 <- "0 ng/ml"
tag3@meta.data$phenotype <- "HTAP"
tag3@meta.data$patient <- patient[3]
tag3@meta.data$sample <- tags[3]

tag3 <- SCTransform(object = tag3, verbose = FALSE)
tag3 <- RunPCA(tag3, verbose = FALSE)
tag3 <- RunUMAP(object = tag3, dims = 1:30)
tag3 <- FindNeighbors(object = tag3, do.plot=TRUE, dims = 1:30)
tag3 <- FindClusters(object = tag3, resolution=0.4)
invisible(gc())

DimPlot(object = tag3, label=TRUE)

tag3 <- RenameCells(tag3, add.cell.id = paste0(Lib,"3"))
saveRDS(tag3, paste0(outdir,"/samples/",Lib,"3",".rds"))


### Tag4 analysis

tag4 <- subset(singlet, subset = HTO_maxID == tags[4])
tag4@meta.data$library <- Lib
tag4@meta.data$tag <- paste0(Lib,"4")
tag4@meta.data$BMP9 <- "10 ng/ml"
tag4@meta.data$phenotype <- "HTAP.BMP9"
tag4@meta.data$patient <- patient[4]
tag4@meta.data$sample <- tags[4]

tag4 <- SCTransform(object = tag4, verbose = FALSE)
tag4 <- RunPCA(tag4, verbose = FALSE)
tag4 <- RunUMAP(object = tag4, dims = 1:30)
tag4 <- FindNeighbors(object = tag4, do.plot=TRUE, dims = 1:30)
tag4 <- FindClusters(object = tag4, resolution=0.4)
invisible(gc())

DimPlot(object = tag4, label=TRUE)

tag4 <- RenameCells(tag4, add.cell.id = paste0(Lib,"4"))
saveRDS(tag4, paste0(outdir,"/samples/",Lib,"4",".rds"))

```


## library D
### Qualty Control, hashing demultiplexing

```{r message=FALSE, warning=FALSE, eval=T, fig.height=6, fig.width=16}

Lib <- "D"
patient <-  c("C685116","C685116","H248","H248")
tags <-  c("C685116","C685116.BMP9","H248","H248.BMP9")

data <- Read10X(paste0(workdir,"/data/Lib_",Lib))
hto <- Read10X(paste0(workdir, "/data/hto/Lib_",Lib,"_umi_count"), gene.column=1)
hto = hto[1:4,]
hto[1:4,1:5]
Matrix::rowSums(hto)

test="-1"
colnames(hto) = paste(colnames(hto),test,sep = "")
common <- intersect(colnames(data), colnames(hto))
data <- data[,common]
hto <- hto[,common]
rownames(hto) <- tags

hashing <- CreateSeuratObject(data)
hashing[["HTO"]] <- CreateAssayObject(hto)
hashing <- NormalizeData(object = hashing, assay = "HTO", normalization.method = "CLR")
hashing <- HTODemux(hashing)

mito.genes <- grep(pattern = "^MT-", x = rownames(hashing@assays$RNA), value = TRUE)
ribo.genes <- grep(pattern = "^RP[SL]", x = rownames(hashing@assays$RNA), value = TRUE)
dropouts <- Matrix::colSums(hashing@assays$RNA@data == 0)/nrow(hashing@assays$RNA)
percent.mito <- Matrix::colSums(hashing@assays$RNA[mito.genes, ])/Matrix::colSums(hashing@assays$RNA)
percent.ribo <- Matrix::colSums(hashing@assays$RNA[ribo.genes, ])/Matrix::colSums(hashing@assays$RNA)
hashing[['percent.mito']] <- percent.mito
hashing[['percent.ribo']] <- percent.ribo
hashing[['dropouts']] <- dropouts
VlnPlot(hashing, features = c("nFeature_RNA", "nCount_RNA","dropouts","percent.ribo","percent.mito"), ncol=5)
dim(hashing@assays$RNA)

hashing <- subset(hashing, subset = percent.mito < 0.15)
dim(hashing@assays$RNA)
VlnPlot(hashing, features = c("nFeature_RNA", "nCount_RNA","dropouts","percent.ribo","percent.mito"), ncol=5)

print(table(hashing@meta.data$HTO_classification.global))
hashing <- SetIdent(hashing, value = "HTO_maxID")

HTOHeatmap(hashing)

```

#### Singlet only

```{r message=FALSE, warning=FALSE, eval=T, fig.height=6, fig.width=12 }

hashing <- SetIdent(hashing, value = "HTO_classification.global")
singlet <- subset(hashing, idents = "Singlet")
table(singlet@meta.data$HTO_maxID)

```


#### Tag analysis

```{r message=FALSE, warning=FALSE, eval=T, fig.height=6, fig.width=8 }

### Tag1 analysis

tag1 <- subset(singlet, subset = HTO_maxID == tags[1])
tag1@meta.data$library <- Lib
tag1@meta.data$tag <- paste0(Lib,"1")
tag1@meta.data$BMP9 <- "0 ng/ml"
tag1@meta.data$phenotype <- "CTRL"
tag1@meta.data$patient <- patient[1]
tag1@meta.data$sample <- tags[1]

tag1 <- SCTransform(object = tag1, verbose = FALSE)
tag1 <- RunPCA(tag1, verbose = FALSE)
tag1 <- RunUMAP(object = tag1, dims = 1:30)
tag1 <- FindNeighbors(object = tag1, do.plot=TRUE, dims = 1:30)
tag1 <- FindClusters(object = tag1, resolution=0.4)
invisible(gc())

DimPlot(object = tag1, label=TRUE)

tag1 <- RenameCells(tag1, add.cell.id = paste0(Lib,"1"))
saveRDS(tag1, paste0(outdir,"/samples/",Lib,"1",".rds"))


### Tag2 analysis

tag2 <- subset(singlet, subset = HTO_maxID == tags[2])
tag2@meta.data$library <- Lib
tag2@meta.data$tag <- paste0(Lib,"2")
tag2@meta.data$BMP9 <- "10 ng/ml"
tag2@meta.data$phenotype <- "CTRL.BMP9"
tag2@meta.data$patient <- patient[2]
tag2@meta.data$sample <- tags[2]

tag2 <- SCTransform(object = tag2, verbose = FALSE)
tag2 <- RunPCA(tag2, verbose = FALSE)
tag2 <- RunUMAP(object = tag2, dims = 1:30)
tag2 <- FindNeighbors(object = tag2, do.plot=TRUE, dims = 1:30)
tag2 <- FindClusters(object = tag2, resolution=0.4)
invisible(gc())

DimPlot(object = tag2, label=TRUE)

tag2 <- RenameCells(tag2, add.cell.id = paste0(Lib,"2"))
saveRDS(tag2, paste0(outdir,"/samples/",Lib,"2",".rds"))

### Tag3 analysis

tag3 <- subset(singlet, subset = HTO_maxID == tags[3])
tag3@meta.data$library <- Lib
tag3@meta.data$tag <- paste0(Lib,"3")
tag3@meta.data$BMP9 <- "0 ng/ml"
tag3@meta.data$phenotype <- "HTAP"
tag3@meta.data$patient <- patient[3]
tag3@meta.data$sample <- tags[3]

tag3 <- SCTransform(object = tag3, verbose = FALSE)
tag3 <- RunPCA(tag3, verbose = FALSE)
tag3 <- RunUMAP(object = tag3, dims = 1:30)
tag3 <- FindNeighbors(object = tag3, do.plot=TRUE, dims = 1:30)
tag3 <- FindClusters(object = tag3, resolution=0.4)
invisible(gc())

DimPlot(object = tag3, label=TRUE)

tag3 <- RenameCells(tag3, add.cell.id = paste0(Lib,"3"))
saveRDS(tag3, paste0(outdir,"/samples/",Lib,"3",".rds"))

### Tag4 analysis

tag4 <- subset(singlet, subset = HTO_maxID == tags[4])
tag4@meta.data$library <- Lib
tag4@meta.data$tag <- paste0(Lib,"4")
tag4@meta.data$BMP9 <- "10 ng/ml"
tag4@meta.data$phenotype <- "HTAP.BMP9"
tag4@meta.data$patient <- patient[4]
tag4@meta.data$sample <- tags[4]

tag4 <- SCTransform(object = tag4, verbose = FALSE)
tag4 <- RunPCA(tag4, verbose = FALSE)
tag4 <- RunUMAP(object = tag4, dims = 1:30)
tag4 <- FindNeighbors(object = tag4, do.plot=TRUE, dims = 1:30)
tag4 <- FindClusters(object = tag4, resolution=0.4)
invisible(gc())

DimPlot(object = tag4, label=TRUE)

tag4 <- RenameCells(tag4, add.cell.id = paste0(Lib,"4"))
saveRDS(tag4, paste0(outdir,"/samples/",Lib,"4",".rds"))

```



## library E
### Qualty Control, hashing demultiplexing

```{r message=FALSE, warning=FALSE, eval=T, fig.height=6, fig.width=16}

Lib <- "E"
patient <-  c("C676136","C676136","H250","H250")
tags <-  c("C676136","C676136.BMP9","H250","H250.BMP9")

data <- Read10X(paste0(workdir,"/data/Lib_",Lib))
hto <- Read10X(paste0(workdir, "/data/hto/Lib_",Lib,"_umi_count"), gene.column=1)
hto = hto[1:4,]
hto[1:4,1:5]
Matrix::rowSums(hto)

test="-1"
colnames(hto) = paste(colnames(hto),test,sep = "")
common <- intersect(colnames(data), colnames(hto))
data <- data[,common]
hto <- hto[,common]
rownames(hto) <- tags

hashing <- CreateSeuratObject(data)
hashing[["HTO"]] <- CreateAssayObject(hto)
hashing <- NormalizeData(object = hashing, assay = "HTO", normalization.method = "CLR")
hashing <- HTODemux(hashing)

mito.genes <- grep(pattern = "^MT-", x = rownames(hashing@assays$RNA), value = TRUE)
ribo.genes <- grep(pattern = "^RP[SL]", x = rownames(hashing@assays$RNA), value = TRUE)
dropouts <- Matrix::colSums(hashing@assays$RNA@data == 0)/nrow(hashing@assays$RNA)
percent.mito <- Matrix::colSums(hashing@assays$RNA[mito.genes, ])/Matrix::colSums(hashing@assays$RNA)
percent.ribo <- Matrix::colSums(hashing@assays$RNA[ribo.genes, ])/Matrix::colSums(hashing@assays$RNA)
hashing[['percent.mito']] <- percent.mito
hashing[['percent.ribo']] <- percent.ribo
hashing[['dropouts']] <- dropouts
VlnPlot(hashing, features = c("nFeature_RNA", "nCount_RNA","dropouts","percent.ribo","percent.mito"), ncol=5)
dim(hashing@assays$RNA)

hashing <- subset(hashing, subset = percent.mito < 0.15)
dim(hashing@assays$RNA)
VlnPlot(hashing, features = c("nFeature_RNA", "nCount_RNA","dropouts","percent.ribo","percent.mito"), ncol=5)

print(table(hashing@meta.data$HTO_classification.global))
hashing <- SetIdent(hashing, value = "HTO_maxID")

HTOHeatmap(hashing)


```

#### Singlet only

```{r message=FALSE, warning=FALSE, eval=T, fig.height=6, fig.width=12 }

hashing <- SetIdent(hashing, value = "HTO_classification.global")
singlet <- subset(hashing, idents = "Singlet")
table(singlet@meta.data$HTO_maxID)

```


#### Tag analysis

```{r message=FALSE, warning=FALSE, eval=T, fig.height=6, fig.width=8 }

### Tag1 analysis

tag1 <- subset(singlet, subset = HTO_maxID == tags[1])
tag1@meta.data$library <- Lib
tag1@meta.data$tag <- paste0(Lib,"1")
tag1@meta.data$BMP9 <- "0 ng/ml"
tag1@meta.data$phenotype <- "CTRL"
tag1@meta.data$patient <- patient[1]
tag1@meta.data$sample <- tags[1]

tag1 <- SCTransform(object = tag1, verbose = FALSE)
tag1 <- RunPCA(tag1, verbose = FALSE)
tag1 <- RunUMAP(object = tag1, dims = 1:30)
tag1 <- FindNeighbors(object = tag1, do.plot=TRUE, dims = 1:30)
tag1 <- FindClusters(object = tag1, resolution=0.4)
invisible(gc())

DimPlot(object = tag1, label=TRUE)

tag1 <- RenameCells(tag1, add.cell.id = paste0(Lib,"1"))
saveRDS(tag1, paste0(outdir,"/samples/",Lib,"1",".rds"))

### Tag2 analysis

tag2 <- subset(singlet, subset = HTO_maxID == tags[2])
tag2@meta.data$library <- Lib
tag2@meta.data$tag <- paste0(Lib,"2")
tag2@meta.data$BMP9 <- "10 ng/ml"
tag2@meta.data$phenotype <- "CTRL.BMP9"
tag2@meta.data$patient <- patient[2]
tag2@meta.data$sample <- tags[2]

tag2 <- SCTransform(object = tag2, verbose = FALSE)
tag2 <- RunPCA(tag2, verbose = FALSE)
tag2 <- RunUMAP(object = tag2, dims = 1:30)
tag2 <- FindNeighbors(object = tag2, do.plot=TRUE, dims = 1:30)
tag2 <- FindClusters(object = tag2, resolution=0.4)
invisible(gc())

DimPlot(object = tag2, label=TRUE)

tag2 <- RenameCells(tag2, add.cell.id = paste0(Lib,"2"))
saveRDS(tag2, paste0(outdir,"/samples/",Lib,"2",".rds"))


### Tag3 analysis

tag3 <- subset(singlet, subset = HTO_maxID == tags[3])
tag3@meta.data$library <- Lib
tag3@meta.data$tag <- paste0(Lib,"3")
tag3@meta.data$BMP9 <- "0 ng/ml"
tag3@meta.data$phenotype <- "HTAP"
tag3@meta.data$patient <- patient[3]
tag3@meta.data$sample <- tags[3]

tag3 <- SCTransform(object = tag3, verbose = FALSE)
tag3 <- RunPCA(tag3, verbose = FALSE)
tag3 <- RunUMAP(object = tag3, dims = 1:30)
tag3 <- FindNeighbors(object = tag3, do.plot=TRUE, dims = 1:30)
tag3 <- FindClusters(object = tag3, resolution=0.4)
invisible(gc())

DimPlot(object = tag3, label=TRUE)

tag3 <- RenameCells(tag3, add.cell.id = paste0(Lib,"3"))
saveRDS(tag3, paste0(outdir,"/samples/",Lib,"3",".rds"))

### Tag4 analysis

tag4 <- subset(singlet, subset = HTO_maxID == tags[4])
tag4@meta.data$library <- Lib
tag4@meta.data$tag <- paste0(Lib,"4")
tag4@meta.data$BMP9 <- "10 ng/ml"
tag4@meta.data$phenotype <- "HTAP.BMP9"
tag4@meta.data$patient <- patient[4]
tag4@meta.data$sample <- tags[4]

tag4 <- SCTransform(object = tag4, verbose = FALSE)
tag4 <- RunPCA(tag4, verbose = FALSE)
tag4 <- RunUMAP(object = tag4, dims = 1:30)
tag4 <- FindNeighbors(object = tag4, do.plot=TRUE, dims = 1:30)
tag4 <- FindClusters(object = tag4, resolution=0.4)
invisible(gc())

DimPlot(object = tag4, label=TRUE)

tag4 <- RenameCells(tag4, add.cell.id = paste0(Lib,"4"))
saveRDS(tag4, paste0(outdir,"/samples/",Lib,"4",".rds"))

```



