---
title: "10X HTAP BMP9 - Analysis"
author: '[Yvon Mbouamboua](mbouamboua@ipmc.cnrs.fr)'
output:
  html_document:
    theme: united
    df_print: kable
  html_notebook:
    code_folding: none
    theme: united
    toc: yes
    toc_depth: 3
    toc_float: yes
  pdf_document:
    toc: yes
    toc_depth: '3'
date: 'Compiled: `r Sys.Date()`'
---

```{r setup, include=TRUE, warning=FALSE}
rm(list = ls())
# Chunck options
knitr::opts_chunk$set(echo = TRUE, eval = TRUE, cache = FALSE, cache.lazy = FALSE, tidy = TRUE, message = FALSE, warning = FALSE)

# Define directories
work.dir <- "/data/data_mbouamboua/10x.htap/bmp9"
data.dir <- paste(work.dir, "outs", "rds", sep = "/")
output.dir <- paste(work.dir, "outs", "markers", sep = "/")

# Setting parameters
the.seed <- 1337L # Computational Parameters

## Memory 
options(future.globals.maxSize = 80000 * 1024^2)
```

# Setting seed and loading R packages

```{r seed_packages}
  library(Seurat)
  library(dplyr)
  library(ggplot2)
  library(ggrepel)
  library(DOSE)
  library(clusterProfiler)
  library(org.Hs.eg.db)
set.seed(the.seed)
```


# Load data

```{r fig.width=16, fig.height=6}
htap <- readRDS(paste(data.dir, "clean.rds", sep = "/"))
invisible(gc())
p1 <- DimPlot(htap, group.by = "ALK1", label.color = "white") + DarkTheme() 
p2 <- DimPlot(htap, group.by = "condition") + DarkTheme()
p3 <- DimPlot(htap, group.by = "sample") + DarkTheme()
p1+p2+p3
```

# DE ALK1

```{r}
res.dir <- paste(work.dir, "outs", "markers", "DE_ALK1", sep = "/")
Seurat::Idents(htap) <- "ALK1"
test.use = "wilcox"
DefaultAssay(htap) <- "RNA"
htap <- NormalizeData(htap)
Idents(htap) <- "ALK1"
clusters <- sort(as.vector(unique(htap@active.ident)))
clusters
nb <- length(clusters)

## Find all markers per cluster
for(i in 1:nb){
  print(clusters[i])
  Idents(htap) <- "ALK1"
  de_genes <- FindMarkers(htap, ident.1 = clusters[i], ident.2 = NULL, only.pos = FALSE, test.use = test.use, assay = "RNA", slot = "data")
  de_genes$cluster <- replicate(nrow(de_genes),clusters[i])
  de_genes$gene <- rownames(de_genes)
  de_genes$test.use <- replicate(nrow(de_genes),paste0(test.use))
  write.table(de_genes,file=paste(res.dir, "/" ,clusters[i], "_DE_", test.use,".txt",sep=""),sep="\t",quote=FALSE,row.names=TRUE,col.names=TRUE)
}

## Save DE results in a joined table
joined_res <- list.files(paste(res.dir), pattern="*.txt", full.names=TRUE)
joined_res <- plyr::ldply(joined_res, read.delim)
write.table(joined_res,file=paste(res.dir, "/", "DE_ALK1_",test.use,".tsv",sep=""),sep="\t",quote=FALSE,row.names=FALSE,col.names=TRUE)

```

### Dotplot
```{r}
top = joined_res %>% dplyr::group_by(cluster) %>% do(head(., n=10))
DotPlot(object = htap, assay = "RNA", features=unique(top$gene), dot.scale = 4) +
  theme_bw(base_line_size = 0) + coord_flip()

DotPlot(object = htap, assay = "RNA", features=rownames(head(markers_50[order(markers_50$avg_log2FC, decreasing = T),],50)), dot.scale = 4) +
  theme_bw(base_line_size = 0) + coord_flip()
```


```{r warning=FALSE, fig.width=3, fig.height=8}
de_genes <- read.delim(paste(res.dir,"DE_ALK1_wilcox.tsv", sep = "/"), header = T, sep = "\t")
de_genes <- de_genes[de_genes$avg_log2FC > 0,]
top = de_genes %>% dplyr::group_by(cluster) %>% do(head(., n=10))

#kable(head(neg_markers,10))
markers_50 <- head(de_genes,50)
labels <- c(expression("PMECs ALK1"^italic("Low")),
            expression("PMECs ALK1"^italic("High")),
            expression("PMECs ALK1"^italic("Low")*"+"*"BMP9"),
            expression("PMECs ALK1"^italic("High")*"+"*"BMP9")
)

DotPlot(object = htap, assay = "RNA", features=unique(top$gene), dot.scale = 4) +
  theme_bw(base_line_size = 0) + 
  #theme_classic() +
  labs(title = "", y = "Cells", x = "Genes") +
  coord_flip() +
  geom_point(aes(size=pct.exp), shape = 21, colour="black", stroke=0.5) +
  scale_y_discrete(labels = labels, position = "right") +
  scale_colour_gradient2(low = "#0570b0", mid = "white", high = "#d7301f") +
  theme(axis.title = element_blank(),
        text = element_text(size = 10),
        axis.text.x = element_text(angle = 45, 
                                   hjust = 0,
                                   vjust = 0.2,
                                   color = "black"),
        axis.text.y = element_text(color = "black"),
        legend.key.height = unit(0.3, 'cm'),
        legend.key.width = unit(0.3, 'cm')) +
  guides(colour = guide_colorbar(title = "avg.exp", 
                                 order = 1), 
         size = guide_legend(title = "%cell.exp", 
                             order = 2)) 

```

# GSE ALK1 High

```{r message=FALSE, warning=FALSE, echo=FALSE, eval=T, fig.height=10, fig.width=8}

res.dir <- paste(work.dir, "outs", "gse", sep = "/")
markers <- read.delim(paste(output.dir, "DE_ALK1", "DE_ALK1_wilcox.tsv", sep = "/"), header = T, sep = "\t")
markers <- subset(markers,cluster == "ALK1 High")

original_gene_list <- markers$avg_log2FC
names(original_gene_list) <- markers$gene
gene_list<-na.omit(original_gene_list)
gene_list = sort(gene_list, decreasing = TRUE)

gse <- gseGO(geneList=gene_list, 
             ont = "BP", 
             keyType = "SYMBOL", 
             #nPerm = 10000, 
             minGSSize = 3, 
             maxGSSize = 800, 
             OrgDb = org.Hs.eg.db, 
             pAdjustMethod = "none")

saveRDS(gse, paste(res.dir, "gse.ALK1.High.rds", sep = "/"))
GSE_ALK1 <- data.frame(gse@result)
GSE_ALK1$analyse <- rep("ALK1 High", nrow(GSE_ALK1))
write.table(GSE_ALK1,file=paste(res.dir,"Res_GSE_ALK1.High.tsv",sep="/"),sep="\t",quote=FALSE,row.names=TRUE,col.names=TRUE)

```

## Dotplot

```{r fig.height=12, fig.width=8}
require(DOSE)
dotplot(gse, title = "GO biological process - ALK1 High", showCategory = 50) + ggplot2::theme(
        legend.key.height = unit(0.3, 'cm'),
        legend.key.width = unit(0.3, 'cm')) +
  ggplot2::geom_point(shape = 21, colour="black", stroke=0.5) +
  #viridis::scale_colour_viridis(option="magma") +
  #scale_colour_gradient2(low = "#0570b0", mid = "white", high = "#d7301f") +
  guides(size=guide_legend(override.aes=list(shape=21, colour="black", fill="white")))
  
```

## Score and preranked list of GSEA result

```{r}
enrichplot::gseaplot2(gse, geneSetID = 1, title = gse@result$Description[[1]])
enrichplot::gseaplot2(gse, geneSetID = 1, title = gse@result$Description[[30]])
enrichplot::gseaplot2(gse, geneSetID = 1, title = gse@result$Description[[33]])

```



# GSE ALK1 Low

```{r message=FALSE, warning=FALSE, echo=FALSE, eval=T, fig.height=10, fig.width=8}
res.dir <- paste(work.dir, "outs", "gse", sep = "/")
markers <- read.delim(paste(output.dir, "DE_ALK1", "DE_ALK1_wilcox.tsv", sep = "/"), header = T, sep = "\t")
markers <- subset(markers,cluster == "ALK1 Low")

original_gene_list <- markers$avg_log2FC
names(original_gene_list) <- markers$gene
gene_list<-na.omit(original_gene_list)
gene_list = sort(gene_list, decreasing = TRUE)

gse <- gseGO(geneList=gene_list, 
             ont = "BP", 
             keyType = "SYMBOL", 
             #nPerm = 10000, 
             minGSSize = 3, 
             maxGSSize = 800, 
             OrgDb = org.Hs.eg.db, 
             pAdjustMethod = "none")

saveRDS(gse, paste(res.dir, "gse.ALK1.Low.rds", sep = "/"))
GSE_ALK1 <- data.frame(gse@result)
GSE_ALK1$analyse <- rep("ALK1 Low", nrow(GSE_ALK1))
write.table(GSE_ALK1,file=paste(res.dir,"Res_GSE_ALK1.Low.tsv",sep="/"),sep="\t",quote=FALSE,row.names=TRUE,col.names=TRUE)

```

## Dotplot

```{r fig.height=12, fig.width=8}
require(DOSE)
dotplot(gse, title = "GO biological process - ALK1 Low", showCategory = 50) + ggplot2::theme(
        legend.key.height = unit(0.3, 'cm'),
        legend.key.width = unit(0.3, 'cm')) +
  ggplot2::geom_point(shape = 21, colour="black", stroke=0.5) +
  #viridis::scale_colour_viridis(option="magma") +
  #scale_colour_gradient2(low = "#0570b0", mid = "white", high = "#d7301f") +
  guides(size=guide_legend(override.aes=list(shape=21, colour="black", fill="white")))
  
```


## Score and preranked list of GSEA result

```{r}
enrichplot::gseaplot2(gse, geneSetID = 1, title = gse@result$Description[[1]])
enrichplot::gseaplot2(gse, geneSetID = 1, title = gse@result$Description[[30]])
enrichplot::gseaplot2(gse, geneSetID = 1, title = gse@result$Description[[33]])

```



# ALK1 High vs ALK1 High+BMP9

## DE
```{r}
Idents(htap) <- "ALK1"
de_genes <- FindMarkers(htap, ident.1 = "ALK1 High", ident.2 = "ALK1 High+BMP9", only.pos = FALSE, test.use = test.use, assay = "RNA", slot = "data")
de_genes$gene <- rownames(de_genes)
de_genes$analyse <- rep("ALK1.High_vs_ALK1.High.BMP9", nrow(de_genes))
write.table(de_genes,file=paste(output.dir, "/DE_ALK1/", "ALK1.High_vs_ALK1.High.BMP9_",test.use,".tsv",sep=""),sep="\t",quote=FALSE,row.names=TRUE,col.names=TRUE)

```

## Volcano plot

```{r }

# de_genes$avg_log2FC <- round(de_genes$avg_log2FC, digits = 2)
# de_genes$p_val_adj <- formatC(de_genes$p_val_adj, format = "e", digits = 2)
# de_genes$p_val <- formatC(de_genes$p_val, format = "e", digits = 2)

# Volcano plot
de_genes$diffexpressed <- "No"
de_genes$diffexpressed[de_genes$avg_log2FC > 0.5 & de_genes$p_val_adj < 0.05] <- "Up"
de_genes$diffexpressed[de_genes$avg_log2FC < -0.5 & de_genes$p_val_adj < 0.05] <- "Down"
de_genes$delabel <- NA
de_genes$delabel[de_genes$diffexpressed != "No"] <- de_genes$gene[de_genes$diffexpressed != "No"]

# Volcano plot
p1 <- ggplot(data=de_genes, 
             aes(x=avg_log2FC, 
                 y=-log10(p_val_adj), 
                 col=diffexpressed, 
                 label=delabel)) +
  geom_text_repel(
    size = 2.5,
    point.padding = 0, # additional padding around each point
    min.segment.length = 0, # draw all line segments
    max.time = 1, max.iter = 1e5, # stop after 1 second, or after 100,000 iterations
    box.padding = 0.1 # additional padding around each text label
  ) +
  geom_point(size = 2.5, 
             alpha = 0.4) + # transparent point
  xlim(-3,3) +
   ylim(c(0, 350)) +
  theme_light()+
  scale_color_manual(values=c("#0570b0", "gray60", "#d7301f")) +
  #scale_color_manual(values=c("firebrick", "gray60", "navy")) +
  geom_vline(xintercept=c(-0.5, 0.5), linetype = "dashed", col="grey60") +
  geom_hline(yintercept=-log10(0.05),  linetype = "dashed", col="grey60")+
  #labs(title = paste0(celltype), x = "", y = "")+
  theme(legend.position = "none",
        legend.key.size = unit(0.3, "cm")) 
print(p1)
```


## Dotplot

```{r warning=FALSE, fig.width=2.2, fig.height=8}
de_genes <- read.delim(paste(output.dir, "DE_ALK1","ALK1.High_vs_ALK1.High.BMP9_wilcox.tsv", sep = "/"), header = T, sep = "\t")
#kable(head(neg_markers,10))
markers_50 <- head(de_genes,50)

labels <- c(expression("PMECs ALK1"^italic("High")),
            expression("PMECs ALK1"^italic("High")*"+BMP9")
            #expression("PMECs ALK1"^italic("High")*"+"*"BMP9"),
            #expression("PMECs ALK1"^italic("low")*"+"*"BMP9")
)

DotPlot(object = htap, assay = "RNA", features = rownames(head(markers_50[order(markers_50$avg_log2FC, decreasing = T),],50)), idents = c("ALK1 High", "ALK1 High+BMP9"), dot.scale = 4) +
  theme_bw(base_line_size = 0) + 
  #theme_classic() +
  labs(title = "", y = "Cells", x = "Genes") +
  coord_flip() +
  geom_point(aes(size=pct.exp), shape = 21, colour="black", stroke=0.5) +
  scale_y_discrete(labels = labels, position = "right") +
  scale_colour_gradient2(low = "#0570b0", mid = "white", high = "#d7301f") +
  theme(axis.title = element_blank(),
        text = element_text(size = 10),
        axis.text.x = element_text(angle = 45, 
                                   hjust = 0,
                                   vjust = 0.2,
                                   color = "black"),
        axis.text.y = element_text(color = "black"),
        legend.key.height = unit(0.3, 'cm'),
        legend.key.width = unit(0.3, 'cm')) +
  guides(colour = guide_colorbar(title = "avg.exp", 
                                 order = 1), 
         size = guide_legend(title = "%cell.exp", 
                             order = 2)) 
```


## GSE

```{r message=FALSE, warning=FALSE, echo=FALSE, eval=T, fig.height=10, fig.width=8}

res.dir <- paste(work.dir, "outs", "gse", sep = "/")

de_genes <- read.delim(paste(output.dir, "DE_ALK1", "ALK1.High_vs_ALK1.High.BMP9_wilcox.tsv", sep = "/"), header = T, sep = "\t")
#de_genes <- de_genes[de_genes$avg_log2FC > 0,]
original_gene_list <- de_genes$avg_log2FC
names(original_gene_list) <- de_genes$gene
gene_list<-na.omit(original_gene_list)
gene_list = sort(gene_list, decreasing = TRUE)

gse <- gseGO(geneList=gene_list, 
             ont = "BP", 
             keyType = "SYMBOL", 
             #nPerm = 10000, 
             minGSSize = 3, 
             maxGSSize = 800, 
             OrgDb = org.Hs.eg.db, 
             pAdjustMethod = "none")

saveRDS(gse, paste(res.dir, "gse.ALK1.High.vs.ALK1.High.BMP9.rds", sep = "/"))
GSE_ALK1 <- data.frame(gse@result)
GSE_ALK1$analyse <- rep("ALK1 High vs ALK1 High+BMP9", nrow(GSE_ALK1))
write.table(GSE_ALK1,file=paste(res.dir,"Res_GSE_ALK1.High_vs_ALK1.High.BMP9.tsv",sep="/"),sep="\t",quote=FALSE,row.names=TRUE,col.names=TRUE)

```


## Dotplot

```{r fig.height=12, fig.width=8}
require(DOSE)
dotplot(gse, title = "GO BP - ALK1 High vs ALK1 HIgh+BMP9", showCategory = 50) +
    ggplot2::theme(
        legend.key.height = unit(0.3, 'cm'),
        legend.key.width = unit(0.3, 'cm')) +
  ggplot2::geom_point(shape = 21, colour="black", stroke=0.5) +
  #viridis::scale_colour_viridis(option="magma") +
  #scale_colour_gradient2(low = "#0570b0", mid = "white", high = "#d7301f") +
  guides(size=guide_legend(override.aes=list(shape=21, colour="black", fill="white")))
  
```


## Score and preranked list of GSEA result

```{r}
enrichplot::gseaplot2(gse, geneSetID = 1, title = gse@result$Description[[1]])
enrichplot::gseaplot2(gse, geneSetID = 1, title = gse@result$Description[[33]])

```


# ALK1 High+BMP9 vs ALK1 High
## DE

```{r}
Idents(htap) <- "ALK1"
de_genes <- FindMarkers(htap, ident.1 = "ALK1 High+BMP9", ident.2 = "ALK1 High", only.pos = FALSE, test.use = test.use, assay = "RNA", slot = "data")
de_genes$gene <- rownames(de_genes)
de_genes$analyse <- rep("ALK1.High.BMP9_vs_ALK1.High", nrow(de_genes))
write.table(de_genes,file=paste(output.dir, "/DE_ALK1/", "ALK1.High.BMP9_vs_ALK1.High_",test.use,".tsv",sep=""),sep="\t",quote=FALSE,row.names=TRUE,col.names=TRUE)

```



```{r warning=FALSE, fig.width=2.2, fig.height=8}
markers_50 <- head(de_genes,50)
labels <- c(expression("PMECs ALK1"^italic("High")),
            expression("PMECs ALK1"^italic("High")*"+BMP9")
            #expression("PMECs ALK1"^italic("High")*"+"*"BMP9"),
            #expression("PMECs ALK1"^italic("low")*"+"*"BMP9")
)

DotPlot(object = htap, assay = "RNA", features=rownames(head(markers_50[order(markers_50$avg_log2FC, decreasing = T),],50)), idents = c("ALK1 High+BMP9", "ALK1 High"), dot.scale = 4) +
    coord_flip() +
theme_bw(base_line_size = 0) + 
  #theme_classic() +
  labs(title = "", y = "Cells", x = "Genes") +
  geom_point(aes(size=pct.exp), shape = 21, colour="black", stroke=0.5) +
  scale_y_discrete(labels = labels, position = "right") +
  #scale_colour_gradient2(low = "#0570b0", mid = "white", high = "#d7301f") +
  theme(axis.title = element_blank(),
        text = element_text(size = 10),
        axis.text.x = element_text(angle = 45, 
                                   hjust = 0,
                                   vjust = 0.2,
                                   color = "black"),
        axis.text.y = element_text(color = "black"),
        legend.key.height = unit(0.3, 'cm'),
        legend.key.width = unit(0.3, 'cm')) +
  guides(colour = guide_colorbar(title = "avg.exp",order = 1), 
         size = guide_legend(title = "%cell.exp",order = 2)) 
```

## GSE

```{r}
res.dir <- paste(work.dir, "outs", "gse", sep = "/")
# GO
original_gene_list <- de_genes$avg_log2FC
names(original_gene_list) <- de_genes$gene
gene_list<-na.omit(original_gene_list)
gene_list = sort(gene_list, decreasing = TRUE)
#keytypes(org.Hs.eg.db)
gse <- gseGO(geneList=gene_list, 
             ont = "BP", 
             keyType = "SYMBOL", 
             #nPerm = 10000, 
             minGSSize = 3, 
             maxGSSize = 800, 
             OrgDb = org.Hs.eg.db, 
             pAdjustMethod = "none")

saveRDS(gse, paste(res.dir, "gse.ALK1.High.bmp9.vs.ALK1.High.rds", sep = "/"))
GSE_ALK1 <- data.frame(gse@result)
GSE_ALK1$analyse <- rep("ALK1 High+BMP9 vs ALK1 High", nrow(GSE_ALK1))
write.table(GSE_ALK1,file=paste(res.dir,"Res_GSE_ALK1.High.BMP9_vs_ALK1.High.tsv",sep="/"),sep="\t",quote=FALSE,row.names=TRUE,col.names=TRUE)
```

## Dotplot

```{r fig.height=12, fig.width=8}
require(DOSE)
dotplot(gse, title = "GO BP - ALK1 High+BMP9 vs ALK1 High", showCategory = 50) +
    ggplot2::theme(
        legend.key.height = unit(0.3, 'cm'),
        legend.key.width = unit(0.3, 'cm')) +
  ggplot2::geom_point(shape = 21, colour="black", stroke=0.5) +
  #viridis::scale_colour_viridis(option="magma") +
  #scale_colour_gradient2(low = "#0570b0", mid = "white", high = "#d7301f") +
  guides(size=guide_legend(override.aes=list(shape=21, colour="black", fill="white")))
  
```

## Score and preranked list of GSEA result

```{r}
enrichplot::gseaplot2(gse, geneSetID = 1, title = gse@result$Description[[1]])

```

# ALK1 Low vs ALK1 Low+BMP9

## DE

```{r}
Idents(htap) <- "ALK1"
de_genes <- FindMarkers(htap, ident.1 = "ALK1 Low", ident.2 = "ALK1 Low+BMP9", only.pos = FALSE, test.use = test.use, assay = "RNA", slot = "data")
de_genes$gene <- rownames(de_genes)
de_genes$analyse <- rep("ALK1.Low_vs_ALK1.Low.BMP9", nrow(de_genes))
write.table(de_genes,file=paste(output.dir, "/DE_ALK1/", "ALK1.Low_vs_ALK1.Low.BMP9_",test.use,".tsv",sep=""),sep="\t",quote=FALSE,row.names=TRUE,col.names=TRUE)

```


## Dotplot

```{r warning=FALSE, fig.width=2.2, fig.height=8}
de_genes <- read.delim(paste(output.dir, "DE_ALK1","ALK1.Low_vs_ALK1.Low.BMP9_wilcox.tsv", sep = "/"), header = T, sep = "\t")
#kable(head(neg_markers,10))
markers_50 <- head(de_genes,50)

labels <- c(expression("PMECs ALK1"^italic("Low")),
            expression("PMECs ALK1"^italic("Low")*"+BMP9")
)

DotPlot(object = htap, assay = "RNA", features = rownames(head(markers_50[order(markers_50$avg_log2FC, decreasing = T),],50)), idents = c("ALK1 Low", "ALK1 Low+BMP9"), dot.scale = 4) +
  theme_bw(base_line_size = 0) + 
  #theme_classic() +
  labs(title = "", y = "Cells", x = "Genes") +
  coord_flip() +
  geom_point(aes(size=pct.exp), shape = 21, colour="black", stroke=0.5) +
  scale_y_discrete(labels = labels, position = "right") +
  scale_colour_gradient2(low = "#0570b0", mid = "white", high = "#d7301f") +
  theme(axis.title = element_blank(),
        text = element_text(size = 10),
        axis.text.x = element_text(angle = 45, 
                                   hjust = 0,
                                   vjust = 0.2,
                                   color = "black"),
        axis.text.y = element_text(color = "black"),
        legend.key.height = unit(0.3, 'cm'),
        legend.key.width = unit(0.3, 'cm')) +
  guides(colour = guide_colorbar(title = "avg.exp", 
                                 order = 1), 
         size = guide_legend(title = "%cell.exp", 
                             order = 2)) 
```


## GSE

```{r}
res.dir <- paste(work.dir, "outs", "gse", sep = "/")
# GO
original_gene_list <- de_genes$avg_log2FC
names(original_gene_list) <- de_genes$gene
gene_list<-na.omit(original_gene_list)
gene_list = sort(gene_list, decreasing = TRUE)
#keytypes(org.Hs.eg.db)
gse <- gseGO(geneList=gene_list, 
             ont = "BP", 
             keyType = "SYMBOL", 
             #nPerm = 10000, 
             minGSSize = 3, 
             maxGSSize = 800, 
             OrgDb = org.Hs.eg.db, 
             pAdjustMethod = "none")

saveRDS(gse, paste(res.dir, "gse.ALK1.Low.vs.ALK1.Low.BMP9.rds", sep = "/"))
GSE_ALK1 <- data.frame(gse@result)
GSE_ALK1$analyse <- rep("ALK1 Low vs ALK1 Low+BMP9", nrow(GSE_ALK1))
write.table(GSE_ALK1,file=paste(res.dir,"Res_GSE_ALK1.Low_vs_ALK1.Low.BMP9.tsv",sep="/"),sep="\t",quote=FALSE,row.names=TRUE,col.names=TRUE)
```

## Dotplot

```{r fig.height=12, fig.width=8}
require(DOSE)
dotplot(gse, title = "GO BP - ALK1 Low vs ALK1 Low+BMP9", showCategory = 50) +
    ggplot2::theme(
        legend.key.height = unit(0.3, 'cm'),
        legend.key.width = unit(0.3, 'cm')) +
  ggplot2::geom_point(shape = 21, colour="black", stroke=0.5) +
  #viridis::scale_colour_viridis(option="magma") +
  #scale_colour_gradient2(low = "#0570b0", mid = "white", high = "#d7301f") +
  guides(size=guide_legend(override.aes=list(shape=21, colour="black", fill="white")))
  
```

## Score and preranked list of GSEA result

```{r}
enrichplot::gseaplot2(gse, geneSetID = 1, title = gse@result$Description[[1]])

```



# ALK1 Low vs ALK1 Low+BMP9

## DE

```{r}
Idents(htap) <- "ALK1"
de_genes <- FindMarkers(htap, ident.1 = "ALK1 Low+BMP9", ident.2 = "ALK1 Low", only.pos = FALSE, test.use = test.use, assay = "RNA", slot = "data")
de_genes$gene <- rownames(de_genes)
de_genes$analyse <- rep("ALK1.Low.BMP9_vs_ALK1.Low", nrow(de_genes))
write.table(de_genes,file=paste(output.dir, "/DE_ALK1/", "ALK1.Low.BMP9_vs_ALK1.Low_",test.use,".tsv",sep=""),sep="\t",quote=FALSE,row.names=TRUE,col.names=TRUE)

```


## Dotplot

```{r warning=FALSE, fig.width=2.2, fig.height=8}
de_genes <- read.delim(paste(output.dir, "DE_ALK1","ALK1.Low.BMP9_vs_ALK1.Low_wilcox.tsv", sep = "/"), header = T, sep = "\t")
#kable(head(neg_markers,10))
markers_50 <- head(de_genes,50)

labels <- c(expression("PMECs ALK1"^italic("Low")),
            expression("PMECs ALK1"^italic("Low")*"+BMP9")
)

DotPlot(object = htap, assay = "RNA", features = rownames(head(markers_50[order(markers_50$avg_log2FC, decreasing = T),],50)), idents = c("ALK1 Low", "ALK1 Low+BMP9"), dot.scale = 4) +
  theme_bw(base_line_size = 0) + 
  #theme_classic() +
  labs(title = "", y = "Cells", x = "Genes") +
  coord_flip() +
  geom_point(aes(size=pct.exp), shape = 21, colour="black", stroke=0.5) +
  scale_y_discrete(labels = labels, position = "right") +
  scale_colour_gradient2(low = "#0570b0", mid = "white", high = "#d7301f") +
  theme(axis.title = element_blank(),
        text = element_text(size = 10),
        axis.text.x = element_text(angle = 45, 
                                   hjust = 0,
                                   vjust = 0.2,
                                   color = "black"),
        axis.text.y = element_text(color = "black"),
        legend.key.height = unit(0.3, 'cm'),
        legend.key.width = unit(0.3, 'cm')) +
  guides(colour = guide_colorbar(title = "avg.exp", 
                                 order = 1), 
         size = guide_legend(title = "%cell.exp", 
                             order = 2)) 
```


## GSE

```{r}
res.dir <- paste(work.dir, "outs", "gse", sep = "/")
# GO
original_gene_list <- de_genes$avg_log2FC
names(original_gene_list) <- de_genes$gene
gene_list<-na.omit(original_gene_list)
gene_list = sort(gene_list, decreasing = TRUE)
#keytypes(org.Hs.eg.db)
gse <- gseGO(geneList=gene_list, 
             ont = "BP", 
             keyType = "SYMBOL", 
             #nPerm = 10000, 
             minGSSize = 3, 
             maxGSSize = 800, 
             OrgDb = org.Hs.eg.db, 
             pAdjustMethod = "none")

saveRDS(gse, paste(res.dir, "gse.ALK1.Low.BMP9.vs.ALK1.Low.rds", sep = "/"))
GSE_ALK1 <- data.frame(gse@result)
GSE_ALK1$analyse <- rep("ALK1 Low+BMP9 vs ALK1 Low", nrow(GSE_ALK1))
write.table(GSE_ALK1,file=paste(res.dir,"Res_GSE_ALK1.Low.BMP9_vs_ALK1.Low.tsv",sep="/"),sep="\t",quote=FALSE,row.names=TRUE,col.names=TRUE)
```

## Dotplot

```{r fig.height=12, fig.width=8}
require(DOSE)
dotplot(gse, title = "GO BP - ALK1 Low+BMP9 vs ALK1 Low", showCategory = 50) +
    ggplot2::theme(
        legend.key.height = unit(0.3, 'cm'),
        legend.key.width = unit(0.3, 'cm')) +
  ggplot2::geom_point(shape = 21, colour="black", stroke=0.5) +
  #viridis::scale_colour_viridis(option="magma") +
  #scale_colour_gradient2(low = "#0570b0", mid = "white", high = "#d7301f") +
  guides(size=guide_legend(override.aes=list(shape=21, colour="black", fill="white")))
  
```

## Score and preranked list of GSEA result

```{r}
enrichplot::gseaplot2(gse, geneSetID = 1, title = gse@result$Description[[1]])

```




# Save GSE results in a joined table

```{r}
res.dir <- paste(work.dir, "outs", "gse", sep = "/")
joined_res <- list.files(paste(res.dir), pattern="*.tsv", full.names=TRUE)
joined_res <- plyr::ldply(joined_res, read.delim)
write.table(joined_res,file=paste(res.dir,"gse_joined_res.tsv",sep="/"),sep="\t",quote=FALSE,row.names=TRUE,col.names=TRUE)

unique(joined_res$analyse)
dim(joined_res)
```

